### 2020-04-20 probabilisticProgrammingMartingales.tex
Consider reduction to values instead of normal forms.
Require terms to be closed for reductions to be defined, and use
"value" insead of "normal form". Also a few little edits.

\section{Language definition}
The language SPCF is a simply-typed lambda calculus with sampling of real numbers from $[0,1]$ and unbounded scoring, following \cite{MakOP20b}. Types and terms are defined as follows, where $r$ is a real, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function and $\Gamma$ is an environment
\begin{align*}
  & \text{types } A, B ::= \textsf{R}  \mid  A \to B \\
  & \text{values } v ::= x  \mid  \lambda x.s  \mid  \underline{r} \\
  & \text{terms } s, t ::= v  \mid  t_1 t_2  \mid  \underline{f}(s_1,\dots ,s_n)  \mid  \tY s  \mid  \tif{s < 0}{t_1}{t_2}  \mid  \tsample  \mid  \tscore(s)
\end{align*}
\begin{align*}
  \frac{}{\Gamma ; x:A \vdash x:A} \qquad
  \frac{\Gamma ; x:A \vdash s : B}{\Gamma \vdash \lambda x.s : A \to B} \qquad
  \frac{}{\underline{r} : \textsf{R}} \qquad
  \frac{\Gamma \vdash s:A \to B \quad \Gamma \vdash t : A}{\Gamma \vdash s t : B} \\ \\
  \frac{\Gamma \vdash s_1:\textsf{R} \dots \Gamma \vdash s_n:\textsf{R}}{\Gamma \vdash \underline{f}(s_1,\dots,s_n) : \textsf{R}} \ f : \mathbb{R}^n \to \mathbb{R} \qquad
  \frac{\Gamma \vdash s : (A \to B) \to (A \to B)}{\Gamma \vdash \tY s : (A \to B)} \\ \\
  \frac{\Gamma \vdash c : \textsf{R} \quad \Gamma \vdash s_1 : A \quad \Gamma \vdash s_2 : A}{\Gamma \vdash \tif{c < 0}{s_1}{s_2} : A} \qquad
  \frac{}{\Gamma \vdash \tsample : \textsf{R}} \qquad
  \frac{\Gamma \vdash s : \textsf{R}}{\Gamma \vdash \tscore (s) : \textsf{R}}
\end{align*}

To define the reduction relation, let evaluation contexts be of the form:
\begin{align*}
  E ::= & \, \cdot \mid E t \mid v E \mid \underline{f}(r_1,\dots ,r_{k-1}, E, s_{k+1}, \dots, s_n) \\ & \mid \tY E \mid \tif{E<0}{s_1}{s_2} \mid \tscore (E)
\end{align*}
then a term reduces if it is formed by substituting a redex in a context i.e.
\begin{align*}
  E[(\lambda x.s) v] & \to E[s[v/x]] \\
  E[\underline f (\underline r_1, \dots , \underline r_n)] & \to E[\underline{f(r_1,\dots,r_n)}] \\
  E[\tY \lambda x. s] & \to E[\lambda z. s[(\tY \lambda x. s)/x] z] \text{ where $z$ is not free in $s$}\\
  E[\tif{\underline r < 0}{s_1}{s_2}] & \to E[s_1] \text{ where }r < 0 \\
  E[\tif{\underline r < 0}{s_1}{s_2}] & \to E[s_2] \text{ where }r \geq 0 \\
  E[\tsample] & \to E[\underline r] \text{ where } r \in [0,1] \\
  E[\tscore(\underline r)] & \to E[\underline r]
\end{align*}

@@

The one-step reduction is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}
\lo{Replace (globally) ``normal form'' by ``value''.}

The result after $n$ steps is then simply $\red^n(M,s) = \overbrace{\red(...\red(}^n M,s)...)$, and the limit $\red^\infty$ can then be defined as a partial function as $\lim_{n \to \infty} \red^n(M,s)$ whenever that sequence becomes constant by reaching a normal form. A term $M$ terminates for a sample sequence $s$ if the limit $\red^\infty(M,s)$ is defined.

The reduction function is measurable, and the set of normal forms is measurable, therefore the set of $s$ such that $M$ terminates at $s$ within $n$ steps is measurable for any $n$, therefore $\{s \mid M \text{ terminates at } s \}$ is measurable. A term $M$ is said to terminate almost surely if $\mu(\{s \mid M \text{ terminates at } s\}) = 1$.

@@

@@

The one-step reduction is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s \circ F(E[R] \to E[N])) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{s(\tsample_i)}],s\mid_{F(E[\underline{s(\tsample_i)}])}) & \text{if } M = E[\tsample_i] \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}

@@

\section{Ranking functions}
Given a probabilistic program (i.e.~a term $M$), in order to construct a supermartingale to prove its a.s.~termination, a function to assign values to each reachable program state is necessary. Let $Rch(M) = \{x \in \Lambda \mid \exists (y_i) : M \to y_0 \to \dots \to y_n \to x\}$, with the $\sigma$-algebra induced as a subset of $\Lambda$, then define a \emph{ranking function on $M$} to be a measurable function $f:\mathit{Rch}(M) \to \mathbb{R}$ such that
\begin{itemize}
    \item $f(N) \geq 0$ for all $N$
    \item $f(E[\tY \lambda x. N] \geq 1+ f(E[\lambda z. N[(\tY \lambda x. N)/x] z]) \text{ where $z$ is not free in $N$}$
    \item $f(E[\tsample]) \geq \int_0^1 f(E[\underline{x}]) \, \mathrm{d}x$

    \lo{I agree that it does not actually matter whether $\mathit{Rch}(M)$ is a measurable subset of $\Lambda$ (the set of SPCF terms).}

    \lo{It is straightforward to see that $E[\cdot] : x \mapsto E[\underline x]$ is a measurable function $\mathbb{R} \to \mathit{Rch}(M)$, using the ``skeletal term'' $\sigma$-algebra on $\mathit{Rch}(M)$, and so, $x \mapsto f(E[\underline{x}])$ is measurable.}

    \lo{I assume you mean $\int_{I} \mathsf{Leb} (\mathrm{d} x) \, f(E[\underline{x}])$.} 
    \akr{Is the meaning of the integral unclear in the context?} 
    \lo{It it -- I was being pedantic. What you write is a Riemann integral, but it coincides with Lebesgue in this case.}

    \akr{I'm not sure whether $Rch(M)$ is a measurable subset of $\Lambda$, so it might not have a natural measure, but I'm pretty sure this doesn't actually matter, and the $\sigma$-algebra at least is well-defined. $Rch(M)$ could fail to be measurable if there are measurable functions $\mathbb R^n \to \mathbb R$ with non-measurable ranges. I don't know whether these exist. They do for the Lebesgue $\sigma$-algebra, but I'm using the Borel ones.}
    \lo{It is worth inserting the preceding (including the example of a measurable function with non-measurable range) as a remark (or as a ``commenting.sty'' note).}
    \lo{For the integral to be well-defined, we need $x \mapsto E[\underline x]$ to be a measurable function $\mathbb{R} \to \mathit{Rch}(M)$.}

@@

Any term for which a ranking function exists is called ``rankable''.

@@

\begin{theorem} \label{rankable implies ast}
  If an SPCF term is rankable, it terminates almost surely.
\end{theorem}
\begin{proof}
Given a term $M$ and a ranking function $f$ for it, define random variables on the probability space $S$ (where $s$ is a random variable) by
\begin{align*}
(M_n,s_n) & = \red^n(M,s) \\
y_0 & = 0 \\
y_{n+1} & = \min \{ k \mid k>y_n, M_k \text{ normal form or of the form } E[\tY N] \}\\
M'_n & = M_{y_n} \\
X_n & = f(M'_n)
\end{align*}
and define a filtration $\mathcal{F}_n = \sigma(M_k, k \leq n)$ (i.e. all the samples used up to step $n$).

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. 
If $M_{y_n}$ isn't already in normal form, then \changed[lo]{$M_{y_n} = E[\tY \lambda x.N]$ for some $E, N$, therefore $M_{y_n+1} = E[\lambda z. N[(\tY \lambda x. N)/x] z]$} and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't normal form, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

@@

\section{Constructing Ranking Functions}
Although rankability implies almost sure termination, the converse does not hold in general. For example,
\begin{equation}
\tif{-\tsample < 0}{\underline{0}}{(\tY \lambda x. x)x}
\end{equation}
terminates in 3 steps with probability 1, but isn't rankable because $(\tY \lambda x. x)x$ is reachable, although that has probability 0. Not only is this counterexample a.s.t, it's positively almost surely terminating i.e.~the expected time to termination is finite.

A ranking function can be constructed under the stronger assumptions that, for every $N$ reachable from $M$, the expected number of $\tY$-reduction steps from $N$ to a normal form is finite. In particular, the expected number of $\tY$-reduction steps from each reachable term is a ranking function.

\akr{I don't know whether you'll see this before I actually write the other bit, but I'm planning on giving a more practical way of constructing ranking functions later using a similar technique, which is why this theorem seems relevant.}
\begin{theorem} \label{minimal}
Given a term $M$, the function $f:Rch(M) \to \mathbb R$ given by $f(N) = \mathbb E [\text{the number of }\tY\text{-reduction steps from }N\text{ to a normal form}]$, if it exists, is the least of all possible ranking functions of $M$.
\end{theorem}
\begin{proof}
Let $f$ be the candidate least ranking function defined above, and suppose $g$ is another ranking function such that $f(N) > g(N)$ for some $N \in Rch(M)$. The difference $g - f$ is then a supermartingale (with the same setup as in the proof of Theorem \ref{rankable implies ast}. The restrictions of $f$ and $g$ to $Rch(N)$ have the same properties assumed of $f$ and $g$, so assume w.l.o.g.~that $N=M$.

@@

\paragraph{}
Even in the case of reasonable simple terms, explicitly constructing a ranking function would be a lot of work, and Theorem \ref{minimal} makes even stronger assumptions than almost sure termination, so it isn't useful for proving it. Take, for example, the term
\begin{align*}
&(\tY \lambda f, n: \\
&\quad \tif{\tsample - 0.5 < 0}{n}{f (n+1)} \\
&) \underline{0}
\end{align*}
which generates a geometric distribution.
    Despite its simplicity, its $Rch$ contains all the terms
\begin{itemize}
    \item $(\tY \lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)}) \underline{i}$
    \item $(\lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)}) (\tY \lambda f, n: tif{\tsample - 0.5 < 0}{n}{f (n+1)}) \underline{i}$
    \item $(\lambda n: \tif{\tsample - 0.5 < 0}{n}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{n}{f (m+1)}) (m+1)}) \underline{i}$
    \item $\tif{\tsample - 0.5 < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\tif{\underline r - 0.5 < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\tif{\underline{r-0.5} < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\underline{i}$
    \item $(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)$.
\end{itemize}
\akr{TODO: make this into a diagram with arrows.}
Even in this simple case, defining a ranking function explicitly is awkward because of the number of cases, although in most cases, because the value need only be greater than or equal to that of the next term in sequence, it suffices to take the ranking function as taking only 3 distinct values.

>>>>

\section{Language definition}
The language SPCF is a simply-typed lambda calculus with sampling of real numbers from $[0,1]$ and unbounded scoring, following \cite{MakOP20b}. Types and terms are defined as follows, where $r$ is a real, $x$ is a variable, $f : \mathbb{R}^n \to \mathbb{R}$ is any measurable function and $\Gamma$ is an environment:
\begin{align*}
  & \text{types } A, B ::= \textsf{R}  \mid  A \to B \\
  & \text{values } v ::= \lambda x.s  \mid  \underline{r} \\
  & \text{terms } s, t ::= v  \mid  x  \mid  t_1 t_2  \mid  \underline{f}(s_1,\dots ,s_n)  \mid  \tY s  \mid  \tif{s < 0}{t_1}{t_2}  \mid  \tsample  \mid  \tscore(s)
\end{align*}
\begin{align*}
  \frac{}{\Gamma ; x:A \vdash x:A} \qquad
  \frac{\Gamma ; x:A \vdash s : B}{\Gamma \vdash \lambda x.s : A \to B} \qquad
  \frac{}{\underline{r} : \textsf{R}} \qquad
  \frac{\Gamma \vdash s:A \to B \quad \Gamma \vdash t : A}{\Gamma \vdash s t : B} \\ \\
  \frac{\Gamma \vdash s_1:\textsf{R} \dots \Gamma \vdash s_n:\textsf{R}}{\Gamma \vdash \underline{f}(s_1,\dots,s_n) : \textsf{R}} \ f : \mathbb{R}^n \to \mathbb{R} \qquad
  \frac{\Gamma \vdash s : (A \to B) \to (A \to B)}{\Gamma \vdash \tY s : (A \to B)} \\ \\
  \frac{\Gamma \vdash c : \textsf{R} \quad \Gamma \vdash s_1 : A \quad \Gamma \vdash s_2 : A}{\Gamma \vdash \tif{c < 0}{s_1}{s_2} : A} \qquad
  \frac{}{\Gamma \vdash \tsample : \textsf{R}} \qquad
  \frac{\Gamma \vdash s : \textsf{R}}{\Gamma \vdash \tscore (s) : \textsf{R}}
\end{align*}

The set of all terms is denoted $\Lambda$, and the set of closed terms is denoted $\Lambda^0$.

\paragraph{}
To define the reduction relation, let evaluation contexts be of the form:
\begin{align*}
  E ::= & \, \cdot \mid E t \mid v E \mid \underline{f}(r_1,\dots ,r_{k-1}, E, s_{k+1}, \dots, s_n) \\ & \mid \tY E \mid \tif{E<0}{s_1}{s_2} \mid \tscore (E)
\end{align*}
then a term reduces if it is formed by substituting a redex in a context i.e.
\begin{align*}
  E[(\lambda x.s) v] & \to E[s[v/x]] \\
  E[\underline f (\underline r_1, \dots , \underline r_n)] & \to E[\underline{f(r_1,\dots,r_n)}] \\
  E[\tY \lambda x. s] & \to E[\lambda z. s[(\tY \lambda x. s)/x] z] \text{ where $z$ is not free in $s$}\\
  E[\tif{\underline r < 0}{s_1}{s_2}] & \to E[s_1] \text{ where }r < 0 \\
  E[\tif{\underline r < 0}{s_1}{s_2}] & \to E[s_2] \text{ where }r \geq 0 \\
  E[\tsample] & \to E[\underline r] \text{ where } r \in [0,1] \\
  E[\tscore(\underline r)] & \to E[\underline r].
\end{align*}

Every closed well-typed term either is a value or reduces to another term.

@@

The one-step reduction is given by the function $\red : \Lambda^0 \times S \to \Lambda^0 \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{\pi_h(s)}],\pi_t(s)) & \text{if } M = E[\tsample] \\
        (M,s) & \text{if } M \text{ is a value}
    \end{array} \right .
\end{equation}

The result after $n$ steps is then simply $\red^n(M,s) = \overbrace{\red(...\red(}^n M,s)...)$, and the limit $\red^\infty$ can then be defined as a partial function as $\lim_{n \to \infty} \red^n(M,s)$ whenever that sequence becomes constant by reaching a value. A term $M$ terminates for a sample sequence $s$ if the limit $\red^\infty(M,s)$ is defined.

The reduction function is measurable, and the set of values is measurable, therefore the set of $s$ such that $M$ terminates at $s$ within $n$ steps is measurable for any $n$, therefore $\{s \mid M \text{ terminates at } s \}$ is measurable. A term $M$ is said to terminate almost surely if $\mu(\{s \mid M \text{ terminates at } s\}) = 1$.

@@

\paragraph{}
Note that the requirement that $M$ be closed is not important for the proofs of the theorems, but is simply required to ensure that $M$ reaches a value, rather than a general normal form like $xy$.

@@

The one-step reduction is given by the function $\red : \Lambda^0 \times S \to \Lambda^0 \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s \circ F(E[R] \to E[N])) & \text{if } M = E[R], R \to N \text{ and } R \neq \tsample \\
        (E[\underline{s(\tsample_i)}],s\mid_{F(E[\underline{s(\tsample_i)}])}) & \text{if } M = E[\tsample_i] \\
        (M,s) & \text{if } M \text{ is a value}
    \end{array} \right .
\end{equation}

@@

\section{Ranking functions}
Given a probabilistic program (i.e.~a closed term $M$), in order to construct a supermartingale to prove its a.s.~termination, a function to assign values to each reachable program state is necessary. Let $Rch(M) = \{x \in \Lambda \mid \exists (y_i) : M \to y_0 \to \dots \to y_n \to x\}$, with the $\sigma$-algebra induced as a subset of $\Lambda$, then define a \emph{ranking function on $M$} to be a measurable function $f:\mathit{Rch}(M) \to \mathbb{R}$ such that
\begin{itemize}
    \item $f(N) \geq 0$ for all $N$
    \item $f(E[\tY \lambda x. N] \geq 1+ f(E[\lambda z. N[(\tY \lambda x. N)/x] z]) \text{ where $z$ is not free in $N$}$
    \item $f(E[\tsample]) \geq \int_I f(E[\underline{x}]) \, \mu_{leb}(\mathrm{d}x)$
    \akr{I'm not sure whether $Rch(M)$ is a measurable subset of $\Lambda$, so it might not have a natural measure, but I'm pretty sure this doesn't actually matter, and the $\sigma$-algebra at least is well-defined. $Rch(M)$ could fail to be measurable if there are measurable functions $\mathbb R^n \to \mathbb R$ with non-measurable ranges. I don't know whether these exist. They do for the Lebesgue $\sigma$-algebra, but I'm using the Borel ones.}
    \lo{It is worth inserting the preceding (including the example of a measurable function with non-measurable range) as a remark (or as a ``commenting.sty'' note).}
    \akr{I'd like to include such a remark, but the only reference I've been able to find for the fact that these functions exist is a Stack Exchange question\footnote{\url{https://math.stackexchange.com/questions/1717282/construction-of-a-borel-measurable-function-mapping-borel-set-to-non-borel-set?rq=1}} which implies that the construction is difficult but doesn't actually say what it is (so I probably won't be able to find it on my own).}

@@

Any closed term for which a ranking function exists is called ``rankable''.

@@

\begin{theorem} \label{rankable implies ast}
  If a closed SPCF term is rankable, it terminates almost surely.
\end{theorem}
\begin{proof}
Given a closed term $M$ and a ranking function $f$ for it, define random variables on the probability space $S$ (where $s$ is a random variable) by
\begin{align*}
(M_n,s_n) & = \red^n(M,s) \\
y_0 & = 0 \\
y_{n+1} & = \min \{ k \mid k>y_n, M_k \text{ a value or of the form } E[\tY N] \}\\
M'_n & = M_{y_n} \\
X_n & = f(M'_n)
\end{align*}
and define a filtration $\mathcal{F}_n = \sigma(M_k, k \leq n)$ (i.e. all the samples used up to step $n$).

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. 
If $M_{y_n}$ isn't already a value, then \changed[lo]{$M_{y_n} = E[\tY \lambda x.N]$ for some $E, N$, therefore $M_{y_n+1} = E[\lambda z. N[(\tY \lambda x. N)/x] z]$} and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't a value, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

@@

\section{Constructing Ranking Functions}
Although rankability implies almost sure termination, the converse does not hold in general. For example,
\begin{equation}
\tif{-\tsample < 0}{\underline{0}}{(\tY \lambda x. x)\underline 0}
\end{equation}
terminates in 3 steps with probability 1, but isn't rankable because $(\tY \lambda x. x) \underline 0$ is reachable, although that has probability 0. Not only is this counterexample a.s.t, it's positively almost surely terminating i.e.~the expected time to termination is finite.

A ranking function can be constructed under the stronger assumptions that, for every $N$ reachable from $M$, the expected number of $\tY$-reduction steps from $N$ to a value is finite. In particular, the expected number of $\tY$-reduction steps from each reachable term is a ranking function.

\akr{I don't know whether you'll see this before I actually write the other bit, but I'm planning on giving a more practical way of constructing ranking functions later using a similar technique, which is why this theorem seems relevant.}
\begin{theorem} \label{minimal}
Given a closed term $M$, the function $f:Rch(M) \to \mathbb R$ given by $f(N) = \mathbb E [\text{the number of }\tY\text{-reduction steps from }N\text{ to a value}]$, if it exists, is the least of all possible ranking functions of $M$.
\end{theorem}
\begin{proof}
Let $f$ be the candidate least ranking function defined above, and suppose $g$ is another ranking function such that $f(N) > g(N)$ for some $N \in Rch(M)$. The difference $g - f$ is then a supermartingale (with the same setup as in the proof of Theorem \ref{rankable implies ast}). The restrictions of $f$ and $g$ to $Rch(N)$ have the same properties assumed of $f$ and $g$, so assume w.l.o.g.~that $N=M$.

@@

\paragraph{}
Even in the case of reasonable simple terms, explicitly constructing a ranking function would be a lot of work, and Theorem \ref{minimal} makes even stronger assumptions than almost sure termination, so it isn't useful for proving it. Take, for example, the term
\begin{align*}
&(\tY \lambda f, n: \\
&\quad \tif{\tsample - 0.5 < 0}{n}{f (n+1)} \\
&) \underline{0}
\end{align*}
which generates a geometric distribution.
    Despite its simplicity, its $Rch$ contains all the terms
\begin{itemize}
    \item $(\tY \lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)}) \underline{i}$
    \item $(\lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)}) (\tY \lambda f, n: \tif{\tsample - 0.5 < 0}{n}{f (n+1)}) \underline{i}$
    \item $(\lambda n: \tif{\tsample - 0.5 < 0}{n}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{n}{f (m+1)}) (m+1)}) \underline{i}$
    \item $\tif{\tsample - 0.5 < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\tif{\underline r - 0.5 < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\tif{\underline{r-0.5} < 0}{\underline{i}}{(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)}$
    \item $\underline{i}$
    \item $(\tY \lambda f, m: \tif{\tsample - 0.5 < 0}{m}{f (m+1)}) (\underline{i}+1)$.
\end{itemize}
\akr{TODO: make this into a diagram with arrows.}
Even in this simple case, defining a ranking function explicitly is awkward because of the number of cases, although in most cases, because the value need only be greater than or equal to that of the next term in sequence, it suffices to take the ranking function as taking only 3 distinct values.
